---
title: "main"
author: "Andrew Nam"
output: html_document
date: "01/22/2023"
---

# Setup

```{r setup, include=FALSE}
# these options here change the formatting of how comments are rendered
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      results = "hold",
                      fig.show = "hold")

# Clear any existing variables
rm(list = ls())
```

```{r}
library(tidyverse)
library(janitor)
library(ggthemes)
library(furrr)  # parallelization support for map
library(ggsignif)
library(emmeans)
library(lmerTest)
library(brms)
library(MuMIn)
library(broom)
library(broom.mixed)
library(latex2exp)
library(png)
library(egg)          # for geom_custom
library(grid)         # functions for dealing with images 
library(magick)  # image processing
library(patchwork)


source("theme_publication.R")
theme_set(theme_Publication())
conditionPalette = c("#F8766D", "#00A9FF", "#C77CFF")
miscPalette = c("#FF781F", "#00BA38")
plan(multisession, workers = 8)  # number of cores to use
```

# Load data

```{r}
df_subjects = read_tsv("../../data/tables/df_subjects.tsv") %>% 
  clean_names()
df_demographics = read_tsv("../../data/tables/df_demographics.tsv") %>% 
  clean_names()
df_timestamps = read_tsv("../../data/tables/df_timestamps.tsv")%>% 
  clean_names()
df_trials = read_tsv("../../data/tables/df_trials.tsv") %>% 
  clean_names()
df_graphs = read_tsv("../../data/tables/df_graphs.tsv") %>% 
  clean_names()
df_posteriors = read_tsv("../../data/tables/df_posteriors.tsv") %>% 
  clean_names() %>% 
  mutate(edge = tolower(edge))
df_optimal = read_tsv("../../data/tables/df_optimal.tsv") %>% 
  clean_names()
df_model_graphs = read_tsv("../../data/tables/model_graphs.tsv") %>% 
  clean_names()
df_model_asymptote = read_tsv("../../data/tables/model_asymptote.tsv") %>% 
  clean_names()

# Device figures
df_images = df_graphs %>% 
  select(graph_type) %>% 
  distinct() %>% 
  mutate(img = map(.x = graph_type,
                   .f = ~ readPNG(str_glue("../../figures/experiment/device_types/{.x}.png"))),
         graph_type = factor(graph_type, c('separate', 'pair', 'line',
                                           'common_cause', 'common_effect', 'skip')),
         # condition doesn't really exist, but it bugs out during plotting if not included
         condition = 'Observation Only',
         condition = condition %>% 
           factor(c('Observation Only', 'Explanation Only', 'Observation + Explanation')),
         name = case_when(graph_type == 'separate' ~ 'Separate',
                          graph_type == 'pair' ~ 'Pair',
                          graph_type == 'line' ~ 'Chain',
                          graph_type == 'common_cause' ~ 'Common Cause',
                          graph_type == 'common_effect' ~ 'Common Effect',
                          graph_type == 'skip' ~ 'Skip',
                          T ~ 'NA')) %>% 
  arrange(graph_type)
```

## Wrangle data

### Demographics
```{r}
df_demographics = df_subjects %>% 
  filter(completed) %>% 
  select(pid) %>% 
  left_join(df_demographics) %>% 
  mutate(across(is.character, ~ case_when(.x == 'DATA_EXPIRED' ~ as.character(NA),
                                          T ~ .x)),
         age = as.numeric(age))
```

### Explanations
```{r}
df_explanations = df_trials %>% 
  select(pid, graph_type, obs_index,
         xnode1 = explanation_node1,
         xnode2 = explanation_node2,
         xactive = explanation_nodes_active,
         xcause = explanation_cause) %>% 
  mutate(obs_index = 1 + obs_index,
         xforward = xnode1 < xnode2,
         xedge = ifelse(xforward, str_glue("{xnode1}{xnode2}"), str_glue("{xnode2}{xnode1}")),
         xedge = tolower(xedge)) 
```


### Responses
```{r}
df_responses = df_subjects %>% 
  filter(completed) %>% 
  select(pid, condition) %>% 
  left_join(df_trials) %>% 
  select(pid, condition, graph_type, obs_index,
         response_ab = edge_ab_response,
         response_ac = edge_ac_response,
         response_bc = edge_bc_response) %>% 
  pivot_longer(starts_with('response'),
               names_to = 'edge',
               values_to = 'human',
               names_prefix = 'response_')

# df_optimal; only keep ones corresponding to participants in df_subjects
df_responses = df_subjects %>% 
  filter(completed) %>% 
  select(pid) %>% 
  left_join(df_optimal) %>% 
  mutate(edge = tolower(edge)) %>% 
  select(pid, graph_type, obs_index, edge, model = response_sample) %>% 
  inner_join(df_responses, .)

df_responses = df_graphs %>% 
  pivot_longer(c(ab, ac, bc),
               names_to = 'edge',
               values_to = 'actual') %>% 
  left_join(df_responses, .)

df_responses = df_responses %>%
  pivot_longer(c(human, model),
               names_to = 'source',
               values_to = 'response') %>% 
  mutate(obs_index = obs_index + 1,
         correct = response == actual,
         graph_type = factor(graph_type, c('separate', 'pair', 'line',
                                           'common_cause', 'common_effect', 'skip')))
```

# About participants

## Number of participants in each condition

```{r}
df_subjects %>% 
  filter(completed) %>% 
  group_by(condition) %>% 
  count()
```

## Demographics

### Age
```{r}
df_demographics %>% 
  summarize(mean = mean(age, na.rm=T),
            sd = sd(age, na.rm=T))
```

### Gender
```{r}
df_demographics %>% 
  group_by(sex) %>% 
  count()
```

### Race
```{r}
df_demographics %>% 
  group_by(ethnicity_simplified) %>% 
  count()
```

# Main plot
```{r}
df_formatted = df_responses %>%
  mutate(condition = case_when(condition == 'activations' ~ 'Observation Only',
                               condition == 'explanations' ~ 'Explanation Only',
                               T ~ 'Observation + Explanation') %>% 
           factor(c('Observation Only', 'Explanation Only', 'Observation + Explanation')),
         correct = as.numeric(correct))

df_asymptote = df_images %>% 
  select(condition, graph_type) %>% 
  mutate(asymptote = c(1, 5/6, 2/3, 2/3, 1, 1/2))
df_asymptote = df_asymptote %>% 
  select(graph_type) %>% 
  crossing(condition = c("Explanation Only", "Observation + Explanation"),
           asymptote = 1) %>% 
  rbind(df_asymptote) %>% 
  mutate(condition = condition %>% 
           factor(c('Observation Only', 'Explanation Only', 'Observation + Explanation')))
```

## Average accuracy

```{r}
df = df_formatted %>% 
  group_by(pid, source, condition, graph_type) %>% 
  summarize(accuracy = mean(correct))

p1 = df %>%
  filter(source == 'human') %>% 
  ggplot(aes(x = condition, y = accuracy, fill = condition)) +
  facet_grid(cols = vars(graph_type)) + 
  geom_point(data = df_asymptote,
             aes(y = asymptote),
             shape = 23,
             color = 'black',
             size = 2.5) +
  stat_summary(fun = "mean",
               geom = "bar",
               color = "black",
               alpha = 0.75) +
  geom_point(alpha = 0.25,
             shape = 21,
             position = position_jitter(width = .2, height = 0),
             size = .65) +
  stat_summary(data = df %>%
                 filter(source == 'model'),
               fun.data = "mean_cl_boot",
               shape = 21,
               size = .6,
               stroke = .5,
               geom = "pointrange") +
  stat_summary(fun.data = "mean_cl_boot",
               size = .5,
               geom = "linerange",
               color = 'black',
               alpha = .75) +
  geom_hline(yintercept = 1/3,
             linetype='dashed') +
  geom_custom(data = df_images,
              mapping = aes(data = img,
                            x = 1.25,
                            y = 1.1),
              grob_fun = function(x) rasterGrob(x,
                                                interpolate = T,
                                                vjust = 0,
                                                hjust = 0,
                                                height = .4)) +
  geom_text(data = df_images,
            mapping = aes(label = name),
            x = 2,
            y = 1.7,
            color = 'black') +
  scale_y_continuous(breaks = seq(0, 1, .25),
                     labels = scales::percent) +
  coord_cartesian(clip = "off",
                  ylim = c(0, 1.1)) +
  scale_fill_manual(name="", 
                    values=conditionPalette) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.line.x = element_blank(),
        strip.text = element_blank(),
        panel.spacing.y = unit(1.3, "cm"),
        plot.margin = margin(t = 1.6, l = .2, r = .2, unit = "cm"),
        panel.grid.major.x = element_blank(),
        legend.position = 'none') +
  labs(x = element_blank(),
       y = 'Accuracy')
p1
```

## Accuracy over time

```{r}
p2 = df_formatted %>% 
  mutate(source = ifelse(source == 'human', 'Human', 'Model')) %>% 
  ggplot(aes(x = obs_index,
             y = correct,
             color = condition,
             shape = source,
             group = interaction(condition, source))) +
  stat_summary(aes(linetype = source),
               fun = "mean",
               geom = "line",
               show.legend = FALSE) +
  stat_summary(fun = "mean",
               geom = "point") +
  geom_hline(yintercept = 1/3,
             linetype='dashed') +
  geom_point(aes(shape = 'Asymptote'),
             x = 0, y = 0,
             alpha = 0) +
  scale_shape_manual(values = c(16, 17, 18),
                     breaks = c('Human', 'Model', 'Asymptote')) +
  coord_cartesian(ylim = c(0, 1), clip = 'off') +
  labs(x = 'Trial', y = 'Accuracy', color = "", shape="") +
  guides(colour = guide_legend(order=1),
         shape = guide_legend(order=2)) +
  scale_x_continuous(breaks = seq(2, 10, 2)) +
  scale_y_continuous(breaks = seq(0, 1, .25),
                     labels = scales::percent) +
  scale_color_manual(name="", 
                     values=conditionPalette) +
  theme(legend.position = 'bottom',
        strip.text = element_blank(),
        plot.margin = margin(t = 0, l = .2, r = .2,  unit = "cm")) +
  facet_grid(cols = vars(graph_type))
p2
```

## Knit
```{r}
p1 / p2 + plot_annotation(tag_levels = 'A')

ggsave('../../figures/main.png', width = 8, height = 4.5, scale = 1.25)
```

# Accuracy over time

```{r}
brm_aot_human = df_responses %>% 
  filter(source == 'human') %>% 
  mutate(correct = as.numeric(correct)) %>% 
  brm(correct ~ condition:obs_index + (1 | pid),
      family = 'bernoulli',
      cores = 4,
      seed = 0,
      file = 'brm/brm_aot_human',
      # file_refit = "on_change",
      data = .)

brm_aot_model = df_responses %>% 
  filter(source == 'model') %>% 
  mutate(correct = as.numeric(correct)) %>% 
  brm(correct ~ condition:obs_index + (1 | pid),
      family = 'bernoulli',
      cores = 4,
      seed = 0,
      file = 'brm/brm_aot_model',
      # file_refit = "on_change",
      data = .)
```

```{r}
df = df_responses %>% 
  select(pid, condition) %>% 
  distinct() %>% 
  crossing(obs_index = c(0, 10)) 

df = predict(brm_aot_model, newdata = df) %>% 
  as_tibble() %>% 
  cbind(df, .)

df %>% 
  filter(obs_index == 0) %>%
  summarize(estimate = mean(Estimate),
            ci_low = quantile(Estimate, .025),
            ci_high = quantile(Estimate, .975)) %>% 
  ungroup() %>% 
  mutate(across(is.numeric, ~round(.x, 3)))

df %>% 
  filter(obs_index == 10) %>% 
  group_by(condition, obs_index) %>% 
  summarize(estimate = mean(Estimate),
            ci_low = quantile(Estimate, .025),
            ci_high = quantile(Estimate, .975)) %>% 
  ungroup() %>% 
  mutate(across(is.numeric, ~round(.x, 3)))
```


# Overall Accuracy

## Conditions only

### Humans
```{r}
df_responses %>% 
  filter(source == 'human') %>% 
  glmer(correct ~ condition + (1 | pid), family = 'binomial', data = .) %>% 
  emmeans(specs = pairwise ~ condition,
          adjust = "none")
```

### Model
```{r}
df_responses %>% 
  filter(source == 'model') %>% 
  glmer(correct ~ condition + (1 | pid), family = 'binomial', data = .) %>% 
  emmeans(specs = pairwise ~ condition,
          adjust = "none")
```

## Conditions + graph types

Note: This analysis was not formally included in the manuscript, but it is visualized
in the main plot. This just provides precise numbers for the figure.

Notes about these results
1. The only contrast (out of 3 conditions x 6 devices = 18 possible) that isn't 
significant in the model is between observations-only and explanations-only in the 
separate graph. All other 17 are predicted to be significant.
2. The difference between explanation-only and both is never significant in humans for 
any of the 6 devices. Does this suggest that humans are treating BOTH as primarily
explanation only?
3. For humans, none of the 3 conditions are significantly different from each other for
the separate device.

```{r}
df_contrasts = df_responses %>% 
  group_by(source, graph_type) %>% 
  nest() %>% 
  mutate(df = map(.x = data,
                  .f = ~ (glmer(correct ~ condition + (1 | pid),
                                family = 'binomial',
                                data = .x) %>% 
                            emmeans(specs = pairwise ~ condition,
                                    adjust = "none"))$contrasts %>% 
                    tidy())) %>% 
  select(-data) %>% 
  unnest() %>% 
  select(source, graph_type, contrast, estimate, std.error, statistic, p.value) %>% 
  mutate(across(is.numeric, ~round(.x, 4)))

df_contrasts %>% 
  view()
```


# Comparing models for both condition

Using the data from participants from the 'both' condition, how well does the model
account for the data when it's given observations and explanations (as was presented
to the participants) or when it's given only explanations?

Limitations

1. Since the model often reaches very low posterior probability (almost 0), if a participant
ever makes that response, the log-likelihood will drop significantly. Since the model will
converge faster given both observations and explanations, this will be a bigger effect for
the 'both' model than the 'explanations only' model.
This is an issue since even a single accidental response from a participant can strongly
favor the 'explanation only' model.

2. Comparing LLH in this way assumes that people maintain beliefs about the edges that
approximate the model posteriors.

```{r}
epsilon = 1e-9

df_llh = df_responses %>% 
  filter(condition == 'both') %>% 
  select(pid, source, graph_type, obs_index, edge, response)

df_llh = df_posteriors %>% 
  pivot_longer(starts_with('p_'),
               names_to = 'response',
               values_to = 'probability',
               names_prefix = 'p_') %>% 
  right_join(df_llh) %>% 
  mutate(probability = ifelse(probability <= epsilon, epsilon, probability),
         logp = log10(probability)) %>% 
  group_by(source, pid, condition, graph_type) %>% 
  summarize(logp = sum(logp)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = condition,
              values_from = logp)

df_llh %>% 
  group_by(source, graph_type) %>% 
  summarize_all(sum)
```

```{r}
# How many participants were favored by the explanation-only model compared to both?
df_llh %>% 
  group_by(source, pid) %>% 
  summarize(both = sum(both),
            explanations_only = sum(explanations_only)) %>% 
  mutate(favor_both = both > explanations_only) %>% 
  group_by(source, favor_both) %>% 
  summarize(n = n()) %>% 
  mutate(frac = round(n / sum(n), 4)) %>% 
  arrange(favor_both)

# How many participants were favored by the explanation-only model compared to both for
# each graph type?
df_llh %>% 
  mutate(favor_both = both > explanations_only) %>% 
  group_by(source, graph_type, favor_both) %>% 
  summarize(n = n()) %>% 
  mutate(frac = round(n / sum(n), 4)) %>% 
  filter(favor_both) %>% 
  arrange(source, frac)
```


```{r}
df_llh %>% 
  filter(source == 'human') %>% 
  mutate(x = both - explanations_only,
         favor_both = x > 0) %>% 
  ggplot(aes(x = x)) +
  geom_histogram(aes(color = favor_both), boundary = 0) +
  # ggplot(aes(x = ba, y = be)) +
  # geom_point(alpha = .2) +
  # geom_abline(slope = 1) +
  facet_wrap(vars(graph_type)) +
  labs(x = "log10 odds-ratio")
```




# Edge changes

How often are edge changes made for each condition?
If an edge change is made, how often is it
incorrect -> correct
incorrect -> incorrect
correct -> incorrect

```{r}
num_sims = 2000

df = df_responses %>% 
  filter(source == 'human') %>% 
  group_by(pid, graph_type, edge) %>% 
  arrange(obs_index) %>% 
  mutate(prev = lag(response, default = 'none')) %>% 
  ungroup() %>% 
  mutate(prev_correct = prev == actual,
         change = response != prev,
         c2i = prev_correct & !correct,
         i2c = !prev_correct & correct,
         i2i = change & !prev_correct & !correct) %>% 
  pivot_longer(c(c2i, i2c),
               names_to = 'change_type',
               values_to = 'changed')

df_edge_changes = tibble(sim = 1:2000) %>% 
  mutate(data = future_map(sim, 
                           ~ df %>% 
                             group_by(condition, change_type) %>% 
                             sample_frac(1, replace=TRUE) %>% 
                             summarize(changed = mean(changed), .groups = 'drop') %>% 
                             pivot_wider(names_from = change_type,
                                         values_from = changed) %>% 
                             mutate(diff = i2c - c2i),
                           .options = furrr_options(seed = T),
                           .progress = T)) %>% 
  unnest() 

df_edge_changes %>% 
  pivot_longer(c(c2i, i2c, diff),
               names_to = 'metric') %>% 
  group_by(condition, metric) %>% 
  summarize(estimate = mean(value),
            ci_low = quantile(value, .025),
            ci_high = quantile(value, .975)) %>% 
  mutate(across(is.numeric, ~ round(.x, 4))) %>% 
  filter(metric != 'diff') %>% 
  mutate(condition =  case_when(condition == 'activations' ~ 'Observation Only',
                               condition == 'explanations' ~ 'Explanation Only',
                               T ~ 'Observation + Explanation') %>% 
           factor(c('Observation Only', 'Explanation Only', 'Observation + Explanation')),
         metric = ifelse(metric == 'c2i', 'Correct to Incorrect', 'Incorrect to Correct')) %>% 
  ggplot(aes(x = metric, y = estimate, fill = condition)) +
  geom_bar(stat = 'identity',
           position = position_dodge(width = .9),
           color = 'black',
           alpha = .75) +
  geom_errorbar(aes(ymin=ci_low, ymax=ci_high),
                width=0,
                position=position_dodge(.9)) + 
  scale_fill_manual(name="", 
                    values=conditionPalette) +
  scale_y_continuous(breaks = seq(0, 0.1, .025),
                     labels = scales::percent) +
  labs(x = element_blank(),
       y = "Proportion of Responses") +
  theme(legend.title = element_blank(),
        panel.grid.major.x = element_blank())

ggsave("../../figures/edge_changes.png", width = 6, height = 4, scale = 1)
```


```{r}
df = df_responses %>% 
  filter(source == 'human') %>% 
  group_by(pid, graph_type, edge) %>% 
  arrange(obs_index) %>% 
  mutate(prev = lag(response, default = 'none')) %>% 
  ungroup() %>% 
  mutate(prev_correct = prev == actual,
         change = response != prev,
         c2i = prev_correct & !correct,
         i2c = !prev_correct & correct,
         i2i = change & !prev_correct & !correct) %>% 
  pivot_longer(c(c2i, i2c, i2i),
               names_to = 'change_type',
               values_to = 'changed')

df %>% 
  filter(change_type == 'c2i') %>% 
  glmer(changed ~ condition + (1 | pid),
        family = 'binomial',
        data = .) %>% 
  emmeans(specs = pairwise ~ condition,
          adjust = "none")
```

```{r}
df %>% 
  group_by(condition, change_type) %>% 
  summarize(mean = mean(changed))
```

# Interpreting explanations
```{r}
df_xp = df_explanations %>%
  left_join(df_responses) %>% 
  filter(condition != 'activations') %>% 
  group_by(pid, source, graph_type, edge) %>% 
  mutate(prev = lag(response, default = 'none')) %>% 
  ungroup() %>% 
  mutate(changed = prev != response)
```


## Explained edge

### Bootstrap
```{r}
num_sims = 2000

df_explained = df_xp %>% 
  filter(edge == xedge) %>% 
  mutate(xdirection = ifelse(xforward, 'forward', 'backward'),
         xalign_prev = case_when(prev == 'none' ~ 'none',
                                 prev == xdirection ~ 'same',
                                 T ~ 'reversed'),
         xalign = case_when(response == 'none' ~ 'none',
                            response == xdirection ~ 'same',
                            T ~ 'reversed')) %>% 
  group_by(source, condition, xcause, xactive, xalign_prev) 

df_explained = tibble(sim = 1:num_sims) %>% 
  mutate(data = future_map(sim, 
                           ~ df_explained %>% 
                             sample_frac(1, replace=TRUE) %>% 
                             group_by(source, condition, xcause, xactive, xalign_prev, xalign) %>% 
                             summarize(n = n(), .groups = 'drop_last') %>% 
                             mutate(total = sum(n),
                                    f = n / total),
                           .options = furrr_options(seed = T),
                           .progress = T)) %>% 
  unnest(data)
```

```{r}
df_explained %>% 
  group_by(sim, source, condition, xcause, xalign) %>% 
  summarize(f = sum(n) / sum(total)) %>% 
  group_by(source, condition, xcause, xalign) %>% 
  summarize(estimate = mean(f),
            ci_low = quantile(f, .025),
            ci_high = quantile(f, .975)) %>% 
  ungroup() %>% 
  filter(!xcause,
         xalign != 'same',
         source == 'human')
```


### Plot
```{r}
df_explained %>% 
  group_by(sim, source, condition, xcause, xactive, xalign) %>% 
  summarize(f = sum(n) / sum(total)) %>% 
  group_by(source, condition, xcause, xalign, xactive) %>% 
  summarize(estimate = mean(f),
            ci_low = quantile(f, .025),
            ci_high = quantile(f, .975)) %>% 
  ungroup() %>% 
  filter(!xcause,
         xalign != 'same') %>% 
  mutate(xactive = ifelse(xactive, "Factual", "Counterfactual") %>%
           factor(c("Factual", 'Counterfactual')),
         source = ifelse(source == 'human', 'Human', 'Model'),
         xalign = ifelse(xalign == 'none', 'No Connection', 'Reverse Connection'),
         condition = ifelse(condition == 'explanations', 'Explanation Only', 'Observation + Explanation')) %>% 
  filter(source == 'Human') %>%
  ggplot(aes(x = xactive,
             y = estimate,
             fill = xalign)) +
  geom_bar(position = position_dodge2(width = .9),
           stat = "identity",
           color = 'black',
           alpha = .75) +
  geom_errorbar(aes(ymin=ci_low, ymax=ci_high),
                width=0,
                position=position_dodge(.9)) + 
  scale_x_discrete(labels=c('Factual'=TeX("Factual$^-$"),
                            'Counterfactual'=TeX("Counterfactual$^-$"))) +
  scale_fill_manual(name="", 
                    values=miscPalette) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~condition) +
  labs(x = element_blank(),
       y = "Proportion of Responses") +
  theme(legend.title = element_blank(),
        panel.grid.major.x = element_blank())

ggsave("../../figures/explained_edge.png", width = 6, height = 4)
```

## Unexplained edges

Create DF involving edges that weren't in the explanation.
For these edges, did the responses point to or away from the nodes in explanation, or
no connection?

### Bootstrap

```{r}
num_samples = 2000

# Assume explanation is X -> Y
df_unexplained = df_explanations %>%
  left_join(df_responses) %>% 
  filter(condition != 'activations',
         edge != xedge) %>% 
  mutate(
    # non-explanation edge with X
    edge1 = case_when(xedge == 'ab' & xforward ~ 'ac', 
                      xedge == 'ab' & !xforward ~ 'bc',
                      xedge == 'ac' & xforward ~ 'ab',
                      xedge == 'ac' & !xforward ~ 'bc',
                      xedge == 'bc' & xforward ~ 'ab',
                      T ~ 'ac'),
    # non-explanation edge with Y
    edge2 = case_when(xedge == 'ab' & xforward ~ 'bc',
                      xedge == 'ab' & !xforward ~ 'ac',
                      xedge == 'ac' & xforward ~ 'bc',
                      xedge == 'ac' & !xforward ~ 'ab',
                      xedge == 'bc' & xforward ~ 'ac',
                      T ~ 'ab')) %>% 
  mutate(edge1 = edge1 == edge,
         edge2 = edge2 == edge) %>% 
  mutate(
    # which node is participant's response pointing to
    pointing_to = case_when(response == 'none' ~ 'none', 
                            response == 'forward' ~ substr(edge, 2, 2),
                            T ~ substr(edge, 1, 1)),
    xnode1 = tolower(xnode1),
    xnode2 = tolower(xnode2),
    # is participant's response pointing to the node in explanation
    direction = case_when(response == 'none' ~ 'none', 
                          edge1 & pointing_to == xnode1 ~ 'to_xnode',
                          edge1 & pointing_to != xnode1 ~ 'from_xnode',
                          edge2 & pointing_to == xnode2 ~ 'to_xnode',
                          edge2 & pointing_to != xnode2 ~ 'from_xnode',
                          T ~ 'error')) %>% 
  select(source, condition, xactive, xcause, edge1, edge2, direction) %>% 
  mutate(edge = ifelse(edge1, 'edge1', 'edge2'))

df_unexplained = tibble(sim = 1:num_sims) %>% 
  mutate(data = future_map(sim, 
                           ~ df_unexplained %>% 
                             sample_frac(1, replace=TRUE) %>% 
                             group_by(source, condition, xcause, xactive, edge, direction) %>% 
                             summarize(n = n(), .groups = 'drop_last') %>% 
                             mutate(total = sum(n),
                                    f = n / total),
                           .options = furrr_options(seed = T),
                           .progress = T)) %>% 
  unnest(data)
```

### Plot

```{r}
df_unexplained %>% 
  filter(!xcause) %>% 
  group_by(sim, source, xcause, edge, direction) %>% 
  summarize(n = sum(n),
            total = sum(total),
            f = n / total) %>% 
  group_by(source, xcause, edge, direction) %>% 
  summarize(estimate = mean(f),
            ci_low = quantile(f, .025),
            ci_high = quantile(f, .975)) %>% 
  ungroup() %>% 
  filter(direction != 'none') %>% 
  mutate(edge = ifelse(edge == 'edge1', "X", "Y"),
         direction = ifelse(direction == 'from_xnode', 'From Component', 'To Component'),
         source = ifelse(source == 'human', 'Human', 'Model')) %>%
  ggplot(aes(x = edge,
             y = estimate,
             fill = direction)) +
  geom_bar(position = position_dodge2(width = .9),
           stat = "identity",
           alpha = .75,
           color = 'black') +
  geom_errorbar(aes(ymin=ci_low, ymax=ci_high),
                width=0,
                position=position_dodge(.9)) + 
  scale_fill_manual(name="", 
                    values=miscPalette) +
  scale_y_continuous(labels = scales::percent) +
  facet_wrap(~source) +
  labs(x = element_blank(),
       y = "Proportion of Responses") +
  theme(legend.title = element_blank(),
        panel.grid.major.x = element_blank())

ggsave("../../figures/unexplained_edge.png", width = 6, height = 4)
```